---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# contdid

<!-- badges: start -->
<!-- badges: end -->

An R package for difference-in-differences with a continuous treatment.

## Installation

You can install the development version of contdid from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("bcallaway11/contdid")

library(contdid)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all("~/Dropbox/BMisc")
devtools::load_all("~/Dropbox/ptetools")
devtools::load_all("~/Dropbox/contdid")
```

## Example

To start with, we will simulate some data, where the continuous treatment `D` has no affect on the outcome.
```{r}
# Simulate data
set.seed(123)
sp <- did::reset.sim()
df <- did::build_sim_dataset(sp)
n <- length(unique(df$id))
D <- runif(n, 0, 1)
# add treatment variable
df$D <- BMisc::time_invariant_to_panel(D, df, "id")
```

The following code illustrates how to estimate `ATT(d)` and `ACRT(d)` using the `cont_did` function, provided by the package.

The interface is basically the same as for the `did` package and for other packages that rely on the `ptetools` backend.  The main additional things that need to be provided to the function are the name of the continuous treatment variable, which should be passed through the `dname` argument.

The `cont_did` function expects the continuous treatment variable to behave in certain ways:

1. It needs to be time-invariant.

2. It should be set to its time-invariant value in pre-treatment periods.  This is just a convention of the package, but, in particular, you should not have the treatment variable coded as being 0 in pre-treatment periods.

3. For units that don't participate in the treatment in any time period, the treatment variable just needs to be time-invariant.  In some applications, e.g., the continuous treatment variable may be defined for units that don't actually participate in the treatment.  In other applications, it may not be defined for units that do not participate in the treatment.  The function behaves the same way in either case.

Next, the other important parameters are `target_parameter`, `aggregation`, and `treatment_type`:

* `target_parameter` can either be "level" or "slope".  If "level", then the function will calculate `ATT` parameters.  If set to be "slope", then the function will calculate `ACRT` parameters---these are causal response parameters that are derivatives of the `ATT` parameters.  Our paper @callaway-goodman-santanna-2024 points out some complications for interpreting these derivative type parameters under the most commonly invoked version of the parallel trends assumption.

* `aggregation` can either by "eventstudy" or "dose".  For "eventstudy", depending on the value of the `target_parameter` argument, the function will provide either the average `ATT` across different event times or the average `ACRT` across different event times.  For "dose", the function will average across all time periods and report average affects across different values of the continuous treatment.  For the "dose" aggregation, results are calculated for both `ATT` and `ACRT` and can be displayed by providing different arguments to plotting functions (see example below).

* `treatment_type` can either be "continuous" or "discrete".  Currently only "continuous" is supported.  In this case, the code proceeds as if the treatment really is continuous.  The estimate are computed nonparametrically using B-splines.  The user can control the number of knots and the degree of the B-splines using the `num_knots` and `degree` arguments.  The defaults are `num_knots=0` and `degree=1` which amounts to estimating `ATT(d)` by estimating a linear model in the continuous treatment among treated units and subtracting the average outcome among the comparison units.

### Dose Aggregation

```{r, warning=FALSE, message=FALSE}
cd_res <- cont_did(
  yname = "Y",
  tname = "period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "slope",
  aggregation = "dose",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res)
ggcont_did(cd_res, type = "att")
ggcont_did(cd_res, type = "acrt")
```


### Event Study Aggregations

An event study aggregation for `ATT`

```{r, warning=FALSE, message=FALSE}
cd_res_es_level <- cont_did(
  yname = "Y",
  tname = "period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "level",
  aggregation = "eventstudy",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res_es_level)
ggcont_did(cd_res_es_level)
```

And an event study aggregation for `ACRT`

```{r, warning=FALSE, message=FALSE}
cd_res_es_slope <- cont_did(
  yname = "Y",
  tname = "period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "slope",
  aggregation = "eventstudy",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res_es_slope)
ggcont_did(cd_res_es_slope)
```