---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# contdid

<!-- badges: start -->
<!-- badges: end -->

An R package for difference-in-differences with a continuous treatment.

## Installation

You can install the development version of contdid from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("bcallaway11/contdid")

library(contdid)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all("~/Dropbox/BMisc")
devtools::load_all("~/Dropbox/ptetools")
devtools::load_all("~/Dropbox/contdid")
```

## Conceptual Setup

Below, we give several examples of how to estimate causal effect parameters using the `cont_did` package.

At a high level, the interface is basically the same as for the `did` package and for other packages that rely on the `ptetools` backend, with only a few pieces of additional information being required.  First, the name of the continuous treatment variable should be passed through the `dname` argument.

The `cont_did` function expects the continuous treatment variable to behave in certain ways:

1. It needs to be time-invariant.

2. It should be set to its time-invariant value in pre-treatment periods.  This is just a convention of the package, but, in particular, you should not have the treatment variable coded as being 0 in pre-treatment periods.

3. For units that don't participate in the treatment in any time period, the treatment variable just needs to be time-invariant.  In some applications, e.g., the continuous treatment variable may be defined for units that don't actually participate in the treatment.  In other applications, it may not be defined for units that do not participate in the treatment.  The function behaves the same way in either case.

Next, the other important parameters are `target_parameter`, `aggregation`, and `treatment_type`:

* `target_parameter` can either be "level" or "slope".  If "level", then the function will calculate `ATT` parameters.  If set to be "slope", then the function will calculate `ACRT` parameters---these are causal response parameters that are derivatives of the `ATT` parameters.  Our paper [Callaway, Goodman-Bacon, and Sant'Anna (2024)](https://arxiv.org/abs/2107.02637) points out some complications for interpreting these derivative type parameters under the most commonly invoked version of the parallel trends assumption.

* `aggregation` can either by "eventstudy" or "dose".  For "eventstudy", depending on the value of the `target_parameter` argument, the function will provide either the average `ATT` across different event times or the average `ACRT` across different event times.  For "dose", the function will average across all time periods and report average affects across different values of the continuous treatment.  For the "dose" aggregation, results are calculated for both `ATT` and `ACRT` and can be displayed by providing different arguments to plotting functions (see example below).

* `treatment_type` can either be "continuous" or "discrete".  Currently only "continuous" is supported.  In this case, the code proceeds as if the treatment really is continuous.  The estimate are computed nonparametrically using B-splines.  The user can control the number of knots and the degree of the B-splines using the `num_knots` and `degree` arguments.  The defaults are `num_knots=0` and `degree=1` which amounts to estimating `ATT(d)` by estimating a linear model in the continuous treatment among treated units and subtracting the average outcome among the comparison units.

## Examples

With a continuous treatment, the underlying building blocks are treatment effects that are local to a particular timing group `g` in a particular time period `t` that experienced a particular value of the treatment `d`.  These treatment affects are relatively high-dimensional, and most applications are likely to involve aggregating/combining these underlying parameters.  We focus on aggregations that (i) average across timing-groups and time periods to given average treatment effect parameters as a function of the dose `d` or (ii) averages across doses and partially across timing group and time periods in order to give event studies.

For the results below, we will simulate some data, where the continuous treatment `D` has no effect on the outcome.
```{r}
# Simulate data
set.seed(1234)
df <- simulate_contdid_data(
  n = 5000,
  num_time_periods = 4,
  num_groups = 4,
  dose_linear_effect = 0,
  dose_quadratic_effect = 0
)
head(df)
```

### Case 1: Dose Aggregation

The following code can be used to estimate the `ATT(d)` and `ACRT(d)` parameters for the continuous treatment `D` using the `cont_did` function.  The `aggregation` argument is set to "dose" and the `target_parameter` argument is set to "level" for `ATT(d)` and "slope" for `ACRT(d)`.

```{r, warning=FALSE, message=FALSE}
cd_res <- cont_did(
  yname = "Y",
  tname = "time_period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "slope",
  aggregation = "dose",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res)
ggcont_did(cd_res, type = "att")
ggcont_did(cd_res, type = "acrt")
```


### Case 2: Event Study Aggregations

Next, we consider event study aggregations.  The first is event study aggregation for `ATT`.  The second is event study aggregation for `ACRT`.

**Event study aggregation for `ATT`:**

Notice that the target parameter is set `level` to target ATT, and the `aggregation` argument is set to `eventstudy`.

```{r, warning=FALSE, message=FALSE}
cd_res_es_level <- cont_did(
  yname = "Y",
  tname = "time_period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "level",
  aggregation = "eventstudy",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res_es_level)
ggcont_did(cd_res_es_level)
```

**Event study aggregation for `ACRT`:**

Relative to the previous code, notice that the target parameter is set `slope` to target ACRT.

```{r, warning=FALSE, message=FALSE}
cd_res_es_slope <- cont_did(
  yname = "Y",
  tname = "time_period",
  idname = "id",
  dname = "D",
  data = df,
  gname = "G",
  target_parameter = "slope",
  aggregation = "eventstudy",
  treatment_type = "continuous",
  control_group = "notyettreated",
  biters = 100,
  cband = TRUE,
  num_knots = 1,
  degree = 3,
)

summary(cd_res_es_slope)
ggcont_did(cd_res_es_slope)
```